Download climatic data from AEMET

### 2020-05-06 Visualization-Contagion/Visualizing-Contagion.Rmd
First round of revisions after invitation to revise and resubmit.

Render animation:
```{r cache=TRUE}
animate(anim1, fps = 2, end_pause = 5)
```

@@

>>>>

Render animation:
```{r cache=TRUE, eval=FALSE}
animate(anim1, fps = 2, end_pause = 5)
```

@@

## LISA analysis of contagion

Here I calculate LISA statistics for `Contagion` on each date:
```{r}
junk <- Contagion %>%
  group_by(Date) %>%
  summarize(Ii =, 
            P_Ii = )
localmoran(Contagion$Contagion[Contagion$Date == "2020-03-13"], listw)[,5]
```

LISA statistic calculated for each cross-section
```{r}
Ii <- Contagion %>% 
  group_by(Date) %>%
  group_modify(~ {(localmoran(.x$Contagion, listw)[,1]) %>%
      enframe()}) %>%
  ungroup()  %>%
  transmute(Date,
            ID_INE = rep(c(1:50), 30),
            Ii = value)
```

LISA's p-value:
```{r}
P_Ii <- Contagion %>% 
  group_by(Date) %>%
  group_modify(~ {(localmoran(.x$Contagion, listw)[,5]) %>%
      enframe()}) %>%
  ungroup() %>%
  transmute(Date,
            ID_INE = rep(c(1:50), 30),
            P_Ii = value)
```

Join the results of the LISA analysis to the dataframe:
```{r}
Contagion <- Contagion %>%
  left_join(Ii, by = c("Date" = "Date", "ID_INE" = "ID_INE")) %>%
  left_join(P_Ii, by = c("Date" = "Date", "ID_INE" = "ID_INE"))
```

Label provinces by whether they are positive or negative, and whether they are significant or not:
```{r}
Contagion <- Contagion %>%
  mutate(LISA = case_when(Ii <= 0 & P_Ii <= 0.05 ~ "Hot Spot",
                          Ii <= 0 & P_Ii > 0.05 ~ "NS (Hot)",
                          Ii > 0 & P_Ii <= 0.05 ~ "Cold Spot",
                          Ii > 0 & P_Ii > 0.05 ~ "NS (Cold)"),
         LISA = factor(LISA, 
                       levels = c("Hot Spot",
                                  "NS (Hot)",
                                  "NS (Cold)",
                                  "Cold Spot"),
                       ordered = TRUE))
```

Create a choropleth plot of the provinces based on the results of the LISA analysis:
```{r}
p2 <- ggplot() + 
  geom_sf(data = Contagion %>% 
            filter(CCAA != "Canarias"), 
          aes(fill = LISA, 
              group = interaction(factor(ID_INE), 
                                  Date))) +
  scale_fill_manual(values = c("Hot Spot" = "red", 
                               "NS (Hot)" = "pink", 
                               "NS (Cold)" = "royalblue", 
                               "Cold Spot" = "royalblue4")) + 
  theme_bw()
```

Create an animation using `gganimate`:
```{r}
anim2 <-  p2 +
  labs(title = "Date: {frame_time}") +
  transition_time(Date)
```

Render animation:
```{r cache=TRUE, eval=FALSE}
animate(anim2, fps = 2, end_pause = 5)
```

## Spatial Markov Chains

I used to have Matlab code for spatial markov chains (SMC). It is gone. Here' I will try to reproduce it but now in `R`.

The first step to implement SMC is to discretize the variable of interest. This can be done in two ways: within period quantiles, and pooled quantiles for the whole time series. According to [Rey and Kang](https://pysal.org/notebooks/explore/giddy/Markov_Based_Methods.html), these two forms of doing the analysis give different interpretations: pooled quantiles are absolute trends, whereas within period quantiles are relative dynamics.

Here, we label each case by their pooled quintile (`q_p`):
```{r}
Contagion <- Contagion %>%
  mutate(q_p = cut(Contagion, 
                   quantile(Contagion, c(0, 0.2, 0.4, 0.6, 0.8, 1)), 
                   labels = c("High Contagion", 
                              "Medium-High Contagion", 
                              "Medium Contagion", 
                              "Medium-Low Contagion", 
                              "Low Contagion"))) %>%
  replace_na(replace = list(q_p = "High Contagion"))
```

To label each case by their within day quintile (`q_d`) we need to summarize the quintiles by date:
```{r}
quintiles_by_date <- Contagion %>%
  st_drop_geometry() %>%
  group_by(Date) %>%
  summarize(x20 = quantile(Contagion, 0.2),
            x40 = quantile(Contagion, 0.4),
            x60 = quantile(Contagion, 0.6),
            x80 = quantile(Contagion, 0.8))
```

Join the quintiles by date to the dataframe:
```{r}
Contagion <- Contagion %>%
  left_join(quintiles_by_date, by = "Date")
```

Label the observations according to their values of `Contagion` and the quintiles by date:
```{r}
Contagion <- Contagion %>%
  mutate(q_d = case_when(Contagion <= x20 ~ "High Contagion",
                         Contagion > x20 & Contagion <= x40 ~ "Medium-High Contagion",
                         Contagion > x40 & Contagion <= x60 ~ "Medium Contagion",
                         Contagion > x60 & Contagion <= x80 ~ "Medium-Low Contagion",
                         Contagion > x80 ~ "Low Contagion"),
         q_d = factor(q_d, 
                      levels = c("High Contagion", 
                              "Medium-High Contagion", 
                              "Medium Contagion", 
                              "Medium-Low Contagion", 
                              "Low Contagion"),
                      ordered = TRUE))
```

### 2020-05-06 covid19env/data_processing/00-Donwload-meteo-data-from-AEMET.Rmd
First round of revisions after invitation to revise and resubmit.

```{r libraries}
library(meteoland)
library(aemet)
library(sf)
library(kableExtra)
library(ggplot2)
```

@@

```{r}
load("/Users/fernandoair/Dropbox/COVID-19/covid19-environmental-correlates/covid19env/data/provinces_spain.RData")
```

@@

```{r}
num_prov <- 5
Tem <- list()
for (i in 1:num_prov){
  Tem[[i]] <- downloadAEMEThistorical(apikey, mydates, station_id=provinces_spain$Meteo_Station[i])[[1]]
  print(i)
  Sys.sleep(2)
}
```

## Example: List information
En la lista Tem se obtienen los valores climático

```{r}
kableExtra::kable(head(Tem[[1]]))
```

>>>>

```{r libraries}
library(meteoland)
#devtools::install_github("SevillaR/aemet")
library(aemet)
library(lubridate)
library(nngeo)
library(sf)
library(kableExtra)
library(tidyverse)
```

@@

```{r}
load("provinces_spain.RData")
```

@@

```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[1:40])
```

Extract meteorological data. First initialize the dataframe:
```{r}
df <- data.frame(indicativo = character(), 
                 long = numeric(), 
                 lat = numeric(), 
                 date = ymd(), 
                 Precipitation = numeric(), 
                 MeanTemperature = numeric(), 
                 MinTemperature = numeric(), 
                 MaxTemperature = numeric(), 
                 WindDirection = numeric(), 
                 WindSpeed = numeric(), 
                 SunshineHours = numeric()) 
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Next set of stations:
```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[41:80])
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Next set of stations:
```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[81:120])
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Next set of stations:
```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[121:160])
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Retrieve the next set of stations:
```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[161:200])
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Retrieve the next set of stations:
```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[201:240])
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Retrieve the next set of stations:
```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[241:280])
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Retrieve the next set of stations:
```{r}
Tem <- downloadAEMEThistorical(apikey, mydates, station_id=stations.sf$indicativo[281:291])
```

Extract the data tables for these stations and bind to the dataframe:
```{r}
coords <- coordinates(Tem) %>% data.frame() %>% rownames_to_column("indicativo")
for(i in 1:nrow(Tem@coords)){df <- rbind(df, 
                                          data.frame(indicativo = rep(coords$indicativo[i], 51), 
                                                     long = rep(coords$long[1], 51),
                                                     lat = rep(coords$lat[1], each = 51),
                                                     date = Tem1@dates,
                                                     Tem@data[[i]]))}
```

Rename dataframe:
```{r}
meteo_data <- df %>% left_join(stations.sf %>% st_drop_geometry() %>% transmute(indicativo, Lon, Lat)) %>%
  select(-c(long, lat))
```

Save to RData file:
```{r}
save(meteo_data, file = "meteo_data.RData")
```
